The impact of cosmic variance on simulating weak lensing surveys

### 2014-09-22 Writeup/paper_v3.tex
Confusion between I and F814W sorted out

Photometric redshifts were determined by~\cite{COSMOS_Photoz_30band}.The accuracy of photometric redshifts for F814W $\le$ 22.5 is $\sigma_{\Delta z} = 0.007(1+z_s)$ and at fainter magnitudes, i.e. F814W $\le$ 24, $\sigma_{\Delta z} = 0.012(1+z_s)$.
The photometric redshift values become noisy beyond $z$ of $1$ for our purposes and the various fits to the galaxies are also not very reliable beyond the apparent magnitude $(m)$ value of 23.5. 
Thus, we will exclude all galaxies that have F814W magnitude greater than 23.5. However, we will use F814W$\le 25.2$ sample in volume-limiting process (cf.~\S\ref{sub:volumelimiting}) to calculate completeness.  We take the $z<1.2$ flux-limited sample to fit redshift distribution models (cf.~\S\ref{sub:overdensities}) and then restrict ourselves to $z\le1$ sample for all further analysis.
% ($S/N \sim 50$ {\bf ?} )
% \rachel{Our catalog has the more up-to-date Ilbert et al. photo-$z$,
%   not the older Mobasher ones.  Need to fix citations. Also, we care
%   about the photo-$z$ accuracy beyond 22.5, so it's important to give
%   some information about it to the depth that we actually use.  Finally, I
%   recommend against mentioning the $S/N$ because the calculation is
%   complicated by the correlated noise, so most people use use
%   magnitude instead of $S/N$.}

@@

\begin{figure}
 \centering
  \includegraphics[width=\columnwidth]{redshift_fluxlimited}
  %\caption{\rachel{You show this figure twice.  Why?  The legend needs to say the name of the distribution, not show the functional form.}}
  \caption{Upper panel: Redshift distribution of flux-limited (F814W$\le 23.5$) sample with bins that are 0.05 wide. Two analytical functions with best fit parameters are plotted over it.\
           Lower panel: Plot of $(1+\delta_{g,\text{1D}}) = N/N_{\text{mod}}$ with each functional form as the model for each redshift bin.}
  \label{fig:redshift_fluxlimited}
\end{figure}

@@

\subsection{Volume limiting}
\label{sub:volumelimiting}
COSMOS is a flux-limited survey and is therefore affected by Malmquist bias, i.e. at higher redshifts, the brighter galaxies are preferentially observed.
Our analysis involves comparing galaxies in different redshift slices to find significant differences in morphology, if any, so with a flux-limited sample, we would be comparing only the bright galaxies at high redshifts with
bright and faint galaxies at low redshifts. For a fair comparison, we must restrict ourselves to bright galaxies at all redshifts and this is acheived by volume-limiting the sample.
%\rachel{Explain why you are doing this.  Before saying {\em how} you
%  generate a volume-limited sample, explain {\em why} you are
%  generating one.}
We generate, using the method below, a volume-limited sample that is complete upto $z=1$ by applying a cut on luminosity such that only galaxies instrinsically brighter than a certain threshold is considered. This threshold is set on $K$-corrected $I$-band magnitudes, denoted as $M_I$.
\rachel{Is it $i$ band, or $I$ band?  Check which one and then be
  self-consistent throughout the text.}\arun{I think they call it
  $I_{814}$, so $I$.}
\rachel{But you are still being ambiguous in sometimes calling it $I$
  and other times F814W.  Pick one.}
Since the parent sample contains fainter galaxies, upto F814W=25.2, we compare the distribution of the F814W=23.5 sample with the samples containing fainter galaxies for high redshift bins, to see where the sample is no longer complete.
% \begin{figure}
%  \centering
%  a) \includegraphics[width=\columnwidth]{volume_limiting_pdf}
%  b) \includegraphics[width=\columnwidth]{volume_limiting_cdf}
%  \label{fig:volume_limiting_dist}
%  \caption{ ... }
% \end{figure}

At $M_I\sim-22.0$, we see the sample is beginning to be biased in the $0.9<z<1.0$ bin due to the flux limit. We obtain 97.84\% completeness in this bin for $M_I$, where completeness is defined as the ratio between the number of galaxies in F814W$\le$23.5 and in F814W$\le$25.2 samples.
Thus the sample $z\le1$ and $M_I<-22$ is at least 97.84\% complete. Fig.~\ref{fig:2Dhist} shows that at $M_I<-22$, we are not affected by flux limit yet.

%We see that the region between $0.85<z<1.0$ is neither overdense nor underdense according to our model and hence is not going to be considered for further analysis.
The region between $0.85<z<1.0$ is only moderately overdense and we do not seem to have as many underdense regions to compare with. It would rather be advantageous to disregard this region
and redo the volume limiting procedure.
We relax our luminosity cut %to $-20.8$
so that the sample is volume-limited \emph{not} until $z=1$ but until $z=0.85$. 
We choose to impose the cut at $M_I=-20.8$, with 95.3\% completeness in the $0.8 < z \le 0.85$ bin. This increases the sample size significantly, from 7,418 galaxies to 11,169. We call this sample \s$1$.

\begin{figure}
  \centering
   \includegraphics[width=\columnwidth]{hist2d_mag_mi}
   \caption{2-D histogram of galaxies in apparent magnitude ($m$) and absolute magnitude ($M_I$) space.}
   \label{fig:2Dhist}
 \end{figure}

@@

Thus, we construct a second volume-limited sample \s$2$, this time by letting the luminosity cut evolve. Starting from $M_I = -20.8$ for the $0.8<z\le0.85$ bin, we add to it $1.23$ times the difference between the
bincenters to obtain the luminosity cut for the other (lower) redshift bins. At lower redshifts, we allow for fainter galaxies that wouldn't have passed the cut in \s$1$ and hence has more galaxies.

@@

%Alternatively, we can impose cuts on stellar mass of the galaxies instead of imposing on the luminosity. 
%If we restrict to galaxies with masses $M$ such that $\log(M/M_\odot) \ge 10.15$, then we have at least {\bf 95.0\% } completeness.
Alternatively, one could get around the problem of considering redshift evolution by imposing cuts on stellar mass instead of absolute luminosity in a particular band. In Fig.~\ref{fig:smf}, we show the stellar mass function (SMF) of our sample for various F814W flux limit.
\cite{Tomczak_SMF} report the SMFs for the ZFOURGE survey, which includes COSMOS. They considered for this work, a single stellar population following a Chabrier IMF~\citep{ChabrierIMF}. 
We plot their SMF for \emph{all} in Fig.~\ref{fig:smf} for
reference. Their SMF is higher than ours since they reach $K_s$-band
$5\sigma$ depth of 24.9. 
As in the case of $M_I$ s, we compare the stellar mass in the F814W$\le$23.5 sample with that of the F814W$\le$25.2 sample. 
The sample $\log(M/M_\odot) > 10.15$ is about 95\% complete in the redshift bin $\left[ 0.75 - 0.85\right]$ and has 10,341 galaxies in total.
Thus, we construct a third volume-limited sample \s$3$ by imposing the stellar mass cut mentioned above. The number of galaxies in redshift slices
are tabulated in Table \ref{table:GalaxyCounts} for all 3 ways of obtaining a volume-limited sample. Stellar-mass limited sample happens to be the smallest one.

@@

There is a minor pitfall with this method. We analyse only those galaxies for which postage stamp images exist.
12\% of galaxies that pass our cuts do not have an associated postage stamp image.
Yet, we use the full F814W$\le$23.5 sample, irrespective of the existence of postage stamps, for identifying overdensities and in the completeness calculation for volume-limiting.
%\rachel{What do you mean we use the full sample  $<23.5$?  For what purpose?  We obviously don't use it for the fits  or shape estimates, so be specific.}
%\arun{Added saying we use the full sample for overdensity calculation and for volume-limiting.}
Postage stamps may not exist because, given the size of the galaxy, the size of the postage stamp we want to draw around it (including enough blank space) intersects with the edge of the CCD.
If all galaxies were the same size, this would be a purely random effect, but in fact bigger galaxies are more likely to get excluded by this cut. 
Typically galaxies that are nearby and intrinsically very bright do not have postage stamps associated with them and this is an effect that is dominant at lower redshifts. 
Our completeness calculation is done at high redshifts and thus we believe that our conclusions are not affected by this bias.

@@

\begin{figure}
 \centering
  \includegraphics[width=\columnwidth]{redshift_vollimited}
  \caption{Upper panel: Redshift distribution of volume-limited ($M_I <  -20.8$) \s$1$ sample with bins that are 0.05 wide. Two analytical functions with best fit parameters are plotted over it.\
           Lower panel: Plot of $(1+\delta_{g,\text{1D}}) = N/N_{\text{mod}}$ with each functional form as the model for each redshift bin.}
  \label{fig:redshift_vollimited}
\end{figure}

>>>>

Photometric redshifts were determined by~\cite{COSMOS_Photoz_30band}.The accuracy of photometric redshifts for $m_\text{F814W}\le$ 22.5 is $\sigma_{\Delta z} = 0.007(1+z_s)$ and at fainter magnitudes, i.e. $m_\text{F814W}\le$ 24, $\sigma_{\Delta z} = 0.012(1+z_s)$.
The photometric redshift values become noisy beyond $z$ of $1$ for our purposes and the various fits to the galaxies are also not very reliable beyond the apparent magnitude $(m)$ value of 23.5. 
Thus, we will exclude all galaxies that have F814W magnitude greater than 23.5. However, we will use $m_\text{F814W}\le 25.2$ sample in volume-limiting process (cf.~\S\ref{sub:volumelimiting}) to calculate completeness.  We take the $z<1.2$ flux-limited sample to fit redshift distribution models (cf.~\S\ref{sub:overdensities}) and then restrict ourselves to $z\le1$ sample for all further analysis.
% ($S/N \sim 50$ {\bf ?} )
% \rachel{Our catalog has the more up-to-date Ilbert et al. photo-$z$,
%   not the older Mobasher ones.  Need to fix citations. Also, we care
%   about the photo-$z$ accuracy beyond 22.5, so it's important to give
%   some information about it to the depth that we actually use.  Finally, I
%   recommend against mentioning the $S/N$ because the calculation is
%   complicated by the correlated noise, so most people use use
%   magnitude instead of $S/N$.}

@@

\begin{figure}
 \centering
  \includegraphics[width=\columnwidth]{redshift_fluxlimited}
  %\caption{\rachel{You show this figure twice.  Why?  The legend needs to say the name of the distribution, not show the functional form.}}
  \caption{Upper panel: Redshift distribution of flux-limited ($m_\text{F814W}\le 23.5$) sample with bins that are 0.05 wide. Two analytical functions with best fit parameters are plotted over it.\
           Lower panel: Plot of $(1+\delta_{g,\text{1D}}) = N/N_{\text{mod}}$ with each functional form as the model for each redshift bin.}
  \label{fig:redshift_fluxlimited}
\end{figure}

@@

\subsection{Volume limiting}
\label{sub:volumelimiting}
COSMOS is a flux-limited survey and is therefore affected by Malmquist bias, i.e. at higher redshifts, the brighter galaxies are preferentially observed.
Our analysis involves comparing galaxies in different redshift slices to find significant differences in morphology, if any, so with a flux-limited sample, we would be comparing only the bright galaxies at high redshifts with
bright and faint galaxies at low redshifts. For a fair comparison, we must restrict ourselves to bright galaxies at all redshifts and this is acheived by volume-limiting the sample.
%\rachel{Explain why you are doing this.  Before saying {\em how} you
%  generate a volume-limited sample, explain {\em why} you are
%  generating one.}
We generate, using the method below, a volume-limited sample that is complete upto $z=1$ by applying a cut on luminosity such that only galaxies instrinsically brighter than a certain threshold is considered. This threshold is set on $K$-corrected $F814W$-band magnitudes, denoted as $M_\text{F814W}$.
\rachel{Is it $i$ band, or $I$ band?  Check which one and then be
  self-consistent throughout the text.}\arun{I think they call it
  $I_{814}$, so $I$.}
\rachel{But you are still being ambiguous in sometimes calling it $I$
  and other times F814W.  Pick one.}
Since the parent sample contains fainter galaxies, upto $m_\text{F814W}=25.2$, we compare the distribution of the $m_\text{F814W}=23.5$ sample with the samples containing fainter galaxies for high redshift bins, to see where the sample is no longer complete.
% \begin{figure}
%  \centering
%  a) \includegraphics[width=\columnwidth]{volume_limiting_pdf}
%  b) \includegraphics[width=\columnwidth]{volume_limiting_cdf}
%  \label{fig:volume_limiting_dist}
%  \caption{ ... }
% \end{figure}

At $M_\text{F814W}\sim-22.0$, we see the sample is beginning to be biased in the $0.9<z<1.0$ bin due to the flux limit. We obtain 97.84\% completeness in this bin for $M_\text{F814W}$, where completeness is defined as the ratio between the number of galaxies in $m_\text{F814W}\le 23.5$ and in $m_\text{F814W}\le25.2$ samples.
Thus the sample $z\le1$ and $M_\text{F814W}<-22$ is at least 97.84\% complete. Fig.~\ref{fig:2Dhist} shows that at $M_\text{F814W}<-22$, we are not affected by flux limit yet.

%We see that the region between $0.85<z<1.0$ is neither overdense nor underdense according to our model and hence is not going to be considered for further analysis.
The region between $0.85<z<1.0$ is only moderately overdense and we do not seem to have as many underdense regions to compare with. It would rather be advantageous to disregard this region
and redo the volume limiting procedure.
We relax our luminosity cut %to $-20.8$
so that the sample is volume-limited \emph{not} until $z=1$ but until $z=0.85$. 
We choose to impose the cut at $M_\text{F814W}=-20.8$, with 95.3\% completeness in the $0.8 < z \le 0.85$ bin. This increases the sample size significantly, from 7,418 galaxies to 11,169. We call this sample \s$1$.

\begin{figure}
  \centering
   \includegraphics[width=\columnwidth]{hist2d_mag_mi}
   \caption{2-D histogram of galaxies in apparent magnitude ($m$) and absolute magnitude ($M_\text{F814W}$) space.}
   \label{fig:2Dhist}
 \end{figure}

@@

Thus, we construct a second volume-limited sample \s$2$, this time by letting the luminosity cut evolve. Starting from $M_\text{F814W} = -20.8$ for the $0.8<z\le0.85$ bin, we add to it $1.23$ times the difference between the
bincenters to obtain the luminosity cut for the other (lower) redshift bins. At lower redshifts, we allow for fainter galaxies that wouldn't have passed the cut in \s$1$ and hence has more galaxies.

@@

%Alternatively, we can impose cuts on stellar mass of the galaxies instead of imposing on the luminosity. 
%If we restrict to galaxies with masses $M$ such that $\log(M/M_\odot) \ge 10.15$, then we have at least {\bf 95.0\% } completeness.
Alternatively, one could get around the problem of considering redshift evolution by imposing cuts on stellar mass instead of absolute luminosity in a particular band. In Fig.~\ref{fig:smf}, we show the stellar mass function (SMF) of our sample for various F814W flux limit.
\cite{Tomczak_SMF} report the SMFs for the ZFOURGE survey, which includes COSMOS. They considered for this work, a single stellar population following a Chabrier IMF~\citep{ChabrierIMF}. 
We plot their SMF for \emph{all} in Fig.~\ref{fig:smf} for
reference. Their SMF is higher than ours since they reach $K_s$-band
$5\sigma$ depth of 24.9. 
As in the case of $M_\text{F814W}$ s, we compare the stellar mass in the $m_\text{F814W}\le23.5$ sample with that of the $m_\text{F814W}\le25.2$ sample. 
The sample $\log(M/M_\odot) > 10.15$ is about 95\% complete in the redshift bin $\left[ 0.75 - 0.85\right]$ and has 10,341 galaxies in total.
Thus, we construct a third volume-limited sample \s$3$ by imposing the stellar mass cut mentioned above. The number of galaxies in redshift slices
are tabulated in Table \ref{table:GalaxyCounts} for all 3 ways of obtaining a volume-limited sample. Stellar-mass limited sample happens to be the smallest one.

@@

There is a minor pitfall with this method. We analyse only those galaxies for which postage stamp images exist.
12\% of galaxies that pass our cuts do not have an associated postage stamp image.
Yet, we use the full $m_\text{F814W}\le23.5$ sample, irrespective of the existence of postage stamps, for identifying overdensities and in the completeness calculation for volume-limiting.
%\rachel{What do you mean we use the full sample  $<23.5$?  For what purpose?  We obviously don't use it for the fits  or shape estimates, so be specific.}
%\arun{Added saying we use the full sample for overdensity calculation and for volume-limiting.}
Postage stamps may not exist because, given the size of the galaxy, the size of the postage stamp we want to draw around it (including enough blank space) intersects with the edge of the CCD.
If all galaxies were the same size, this would be a purely random effect, but in fact bigger galaxies are more likely to get excluded by this cut. 
Typically galaxies that are nearby and intrinsically very bright do not have postage stamps associated with them and this is an effect that is dominant at lower redshifts. 
Our completeness calculation is done at high redshifts and thus we believe that our conclusions are not affected by this bias.

@@

\begin{figure}
 \centering
  \includegraphics[width=\columnwidth]{redshift_vollimited}
  \caption{Upper panel: Redshift distribution of volume-limited ($M_\text{F814W} <  -20.8$) \s$1$ sample with bins that are 0.05 wide. Two analytical functions with best fit parameters are plotted over it.\
           Lower panel: Plot of $(1+\delta_{g,\text{1D}}) = N/N_{\text{mod}}$ with each functional form as the model for each redshift bin.}
  \label{fig:redshift_vollimited}
\end{figure}
